---
title: "STAT 897D Team Project"
author: "Raymond Carl"
date: "6 December 2015"
output: pdf_document
---


```{r, echo=FALSE,message=FALSE}
# Load required packages

library(dplyr)
library(leaps)
library(glmnet)
library(splines)
library(gam)


```



```{r, echo=FALSE, message=FALSE}
#Read in data from CSV file
charity <- read.csv("/Users/raymondcarl/Dropbox/STAT 897D Team Project/charity.csv")

```



```{r, echo=FALSE,message=FALSE}
# Helper functions

# Make predictions from regsusets object
# https://github.com/yufree/democode/blob/master/rml/predict.regsubsets.R
predict.regsubsets = function(object, newdata, id, ...) {
        form  <-  as.formula(~.)
        mat  <-  model.matrix(form, newdata)
        coefi  <-  coef(object, id)
        xvars  <-  names(coefi)
        mat[, xvars] %*% coefi
}

#Center the variables
#http://gastonsanchez.com/blog/how-to/2014/01/15/Center-data-in-R.html
center_apply <- function(x) {
    apply(x, 2, function(y) y - mean(y))
}


```

example code for vectorization
```{r}
charity.t <- charity
charity.t$avhv <- log(charity.t$avhv)
data.train <- charity.t[charity$part=="train",]
x.train <- data.train[,2:21]
c.train <- data.train[,22] # donr
n.train.c <- length(c.train) # 3984
y.train <- data.train[c.train==1,23] # damt for observations with donr=1
n.train.y <- length(y.train) # 1995

data.train <- charity.t[charity$part=="train",]
x.train <- data.train[,2:21]
c.train <- data.train[,22] # donr
n.train.c <- length(c.train) # 3984
y.train <- data.train[c.train==1,23] # damt for observations with donr=1
n.train.y <- length(y.train) # 1995

data.valid <- charity.t[charity$part=="valid",]
x.valid <- data.valid[,2:21]
c.valid <- data.valid[,22] # donr
n.valid.c <- length(c.valid) # 2018
y.valid <- data.valid[c.valid==1,23] # damt for observations with donr=1
n.valid.y <- length(y.valid) # 999

data.test <- charity.t[charity$part=="test",]
n.test <- dim(data.test)[1] # 2007
x.test <- data.test[,2:21]

x.train.mean <- apply(x.train, 2, mean)
x.train.sd <- apply(x.train, 2, sd)
x.train.std <- t((t(x.train)-x.train.mean)/x.train.sd) # standardize to have zero mean and unit sd
apply(x.train.std, 2, mean) # check zero mean
apply(x.train.std, 2, sd) # check unit sd
data.train.std.c <- data.frame(x.train.std, donr=c.train) # to classify donr
data.train.std.y <- data.frame(x.train.std[c.train==1,], damt=y.train) # to predict damt when donr=1

x.valid.std <- t((t(x.valid)-x.train.mean)/x.train.sd) # standardize using training mean and sd
data.valid.std.c <- data.frame(x.valid.std, donr=c.valid) # to classify donr
data.valid.std.y <- data.frame(x.valid.std[c.valid==1,], damt=y.valid) # to predict damt when donr=1

x.test.std <- t((t(x.test)-x.train.mean)/x.train.sd) # standardize using training mean and sd
data.test.std <- data.frame(x.test.std)

```




```{r}
#data transformations (ln and ^2)

transformation_df <- charity[,c(11:13,15:21)]

#log transformation
names(transformation_df) <- paste(names(charity[,c(11:13,15:21)]), "_log")
charity<-data.frame(charity, log(transformation_df))

#squared transformation
names(transformation_df) <- paste(names(charity[,c(11:13,15:21)]), "_sq")

transformation_centered_df <- center_apply(transformation_df)
charity<-data.frame(charity, transformation_centered_df^2)

names(transformation_df) <- paste(names(charity[,c(11:13,15:21)]), "_cube")
charity<-data.frame(charity, transformation_centered_df^3)

```


```{r}


names(charity[,1:27])


```





<h4>
Develop a classification model for the DONR Variable
</h4>


Logistic Regression
** This is just a copy of the example guide that came with the project**
```{r}

model.log1 <- glm(donr ~ reg1 + reg2 + reg3 + reg4 + home + chld + hinc + I(hinc^2) + genf + wrat + avhv + incm + inca + plow + npro + tgif + lgif + rgif + tdon + tlag + agif, data.train.std.c, family=binomial("logit"))

post.valid.log1 <- predict(model.log1, data.valid.std.c, type="response")

profit.log1 <- cumsum(14.5*c.valid[order(post.valid.log1, decreasing=T)]-2)
plot(profit.log1) # see how profits change as more mailings are made
n.mail.valid <- which.max(profit.log1) # number of mailings that maximizes profits
c(n.mail.valid, max(profit.log1)) # report number of mailings and maximum profit
# 1291.0 11642.5

cutoff.log1 <- sort(post.valid.log1, decreasing=T)[n.mail.valid+1] # set cutoff based on n.mail.valid
chat.valid.log1 <- ifelse(post.valid.log1>cutoff.log1, 1, 0) # mail to everyone above the cutoff
table(chat.valid.log1, c.valid) # classification table
#               c.valid
#chat.valid.log1   0   1
#              0 709  18
#              1 310 981
# check n.mail.valid = 310+981 = 1291
# check profit = 14.5*981-2*1291 = 11642.5

# Results

# n.mail Profit  Model
# 1329   11624.5 LDA1
# 1291   11642.5 Log1

# select model.log1 since it has maximum profit in the validation sample

post.test <- predict(model.log1, data.test.std, type="response") # post probs for test data

# Oversampling adjustment for calculating number of mailings for test set

n.mail.valid <- which.max(profit.log1)
tr.rate <- .1 # typical response rate is .1
vr.rate <- .5 # whereas validation response rate is .5
adj.test.1 <- (n.mail.valid/n.valid.c)/(vr.rate/tr.rate) # adjustment for mail yes
adj.test.0 <- ((n.valid.c-n.mail.valid)/n.valid.c)/((1-vr.rate)/(1-tr.rate)) # adjustment for mail no
adj.test <- adj.test.1/(adj.test.1+adj.test.0) # scale into a proportion
n.mail.test <- round(n.test*adj.test, 0) # calculate number of mailings for test set

cutoff.test <- sort(post.test, decreasing=T)[n.mail.test+1] # set cutoff based on n.mail.test
chat.test <- ifelse(post.test>cutoff.test, 1, 0) # mail to everyone above the cutoff
table(chat.test)
#    0    1 
# 1676  331
# based on this model we'll mail to the 331 highest posterior probabilities

# See below for saving chat.test into a file for submission


```



<h4>
Prediction Model for the DAMT Variable
</h4>

Note: The first run did not use the transformed variables.


Linear Regression Variable Selection
```{r}

nvmax_var <- 50

training_data <- filter(charity,part=="train" & donr==1) %>% select(.,-donr)
validation_data <- filter(charity,part=="valid" & donr==1) %>% select(.,-donr)

# Backward selection model
# part and ID variables are excluded
regfit.bwd = regsubsets(damt~. -part -ID, data = training_data, nvmax = nvmax_var, method = "backward")

#Test matrix
test.mat=model.matrix(damt~. -part -ID, data=training_data)

#Run a for loop to get the cross validation error
val.errors=rep(NA,nvmax_var) #create a numbered list of x NA's

for(i in 1:nvmax_var){
  coefi=coef(regfit.bwd,id=i)
  pred=test.mat[,names(coefi)]%*%coefi
  val.errors[i]=mean( (training_data$damt - pred)^2 )
}

val.errors
best_model <-which.min(val.errors)

coef(regfit.bwd,best_model)

mean((predict.regsubsets(regfit.bwd, validation_data,best_model) - validation_data$damt)^2)

```

Lasso and Ridge Regression

Vectorize variables for ridge regression and lasso:

```{r}
x=model.matrix(damt~ reg1 + reg2 + reg3 + reg4 + home + chld + hinc + genf + wrat + avhv + incm + inca + plow + npro + tgif + lgif + rgif + tdon + tlag + agif, data=data.train.std.y)
y=data.train.std.y$damt
x.valid=model.matrix(damt~ reg1 + reg2 + reg3 + reg4 + home + chld + hinc + genf + wrat + avhv + incm + inca + plow + npro + tgif + lgif + rgif + tdon + tlag + agif, data=data.valid.std.y)
y.valid=data.valid.std.y$damt

grid=10^seq(10,-2,length=100)

```

Ridge Regression
```{r}

ridge.mod=glmnet(x,y,alpha=0,lambda=grid)
cv.out=cv.glmnet(x,y,alpha=0)
bestlam=cv.out$lambda.min
bestlam
ridge.pred=predict(ridge.mod,s=bestlam,newx=x.valid)
mean((ridge.pred-y.valid)^2)

```

Lasso 
```{r}

lasso.mod=glmnet(x,y,alpha=1,lambda=grid)
cv.out=cv.glmnet(x,y,alpha=1)
bestlam=cv.out$lambda.min
bestlam
lasso.pred=predict(lasso.mod,s=bestlam,newx=x.valid)
mean((lasso.pred-y.valid)^2)

```

Splines
```{r}


fit <- lm(damt~ reg3 + reg4 + bs(incm._log,knots=c(2,2.3,2.7)) + bs(agif._log,knots=c(2,2.3,2.7)),data=training_data)

pred=predict(fit, newdata=validation_data,se=T)
mean((pred$fit-y.valid)^2)

plot(age,wage,col="gray")
lines(age.grid,pred$fit,lwd=2)
lines(age.grid,pred$fit+2*pred$se ,lty="dashed")
lines(age.grid,pred$fit-2*pred$se ,lty="dashed") 

```

